\documentclass[12pt,reqno]{amsart}
%\documentclass[../Solutions_Introduction_to_Algorithms.tex]{subfiles}
\usepackage{amsmath,amsfonts,amscd,amssymb,epsf,color,enumerate,graphicx,url}
\usepackage{algorithm, algorithmic}
\usepackage{forest, tikz, xcolor}
\usetikzlibrary{matrix, positioning}
\setlength{\oddsidemargin}{-0.2in}%
\setlength{\evensidemargin}{-0.2in}%
\setlength{\textwidth}{6.6in}%
\setlength{\topmargin}{-0.5in}%
 \setlength{\textheight}{9.5in}%
 \definecolor{orange}{rgb}{1,0.5,0}
 \pagestyle{plain}
\linespread{1.3}
\usepackage[small]{caption}
\newcommand{\pa}{\partial}
\newcommand{\va}{\vspace{0.4cm}}
\newcommand{\di}{\displaystyle}
\newcommand{\disp}{\displaystyle}


% turn on \answertrue to show the solution
% turn on \answerfalse to hide the solution
\newif\ifanswer
\answertrue
%\answerfalse



\begin{document}
\noindent {\footnotesize Introduction to Algorithms}\hspace{10.5cm} {\footnotesize Solutions}

\vspace{0.5cm}
\hspace{5.5cm}\textbf{\large Exercises in Section 7.3}
\vspace{0.5cm}

\begin{enumerate}[1.]

\item Why do we analyze the expected running time of a randomized algorithm and not its worst-case running time?

\ifanswer
\noindent {\bf Solution}
The running time of a randomized algorithm is random and does not depend on the input parameters (e.g. arrays). Therefore, the worst-case running time, where we discuss the choice of the input parameters, is meaningless for randomized algorithms.
\vspace{1cm}



\item When $\textsc{Randomized-Quicksort}$ runs, how many calls are made to the random-number generator $\textsc{Random}$ in the worst case? How about in the best case? Give your answer in terms of $\Theta$-notation.

\ifanswer
\noindent {\bf Solution}
Note that $\textsc{Random}$ is called once $\textsc{Random-Partition}$ is called. There are $\Theta(n)$ leaves in the recursion tree, and the number of nodes should be a constant product of the number of leaves. Thus, in both worst case and best case, $\textsc{Random}$ is called $\Theta(n)$ times.
\vspace{1cm}



\end{enumerate}

\end{document}



